---
title: "Lake Level Regime - Visualization"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Lake Level Regime - Visualization}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Lake Level Visualization

Alternate ways of visualizing time series of lake levels and different features
of the lake level regime, e.g., magnitude, duration. Developed for use of
figures in presentations which build on one another to introcude these concepts.
Order of plots is always Pleasant (top), Long (middle), and Plainfield (bottom).
Scenario illustrated here is the no-irrigated-agriculture scenario.

<br>

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSscenarios)
library(dplyr)
library(reshape2)
library(NISTunits)
library(lubridate)
library(ggplot2)
library(extrafont)
library(patchwork)

save_on    <- FALSE
fig_width  <- 8.7
fig_height <- 2

scenarios <- data.frame(scenario = c("no_irr", "cur_irr", "fut_irr"),
                        scenario_name = c("No Irrigated Agriculture", 
                                          "Current Irrigated Agriculture",
                                          "Potential Irrigated Agriculture"),
                        colors = c("steelblue4", "goldenrod1", "darkorange"))

# Lake Levels ------------------------------------------------------------------
MODFLOW <- CSLSdata::MODFLOW %>%
           filter(.data$scenario %in% scenarios$scenario,
                  year(.data$date) >= 1986,
                  year(.data$date) <= 2018) %>%
           mutate(year = year(.data$date),
                  month = month(.data$date),
                  time = .data$year - min(.data$year) + .data$month/12,
                  level = NISTmeterTOft(.data$level_m)) %>%
           left_join(scenarios, by = "scenario") %>%
           select(scenario = .data$scenario_name,
                  sim = .data$sim,
                  lake = .data$lake, 
                  date = .data$date,
                  time = .data$time,
                  level = .data$level)

# Metrics ----------------------------------------------------------------------
MODFLOW_metrics <- CSLSscenarios::MODFLOW_metrics %>%
                   group_by(.data$lake, .data$metric, .data$variable,
                            .data$scenario, .data$series) %>%
                   summarise(base = .data$value[which(.data$sim == 1)],
                             sd = sd(.data$value, na.rm = TRUE),
                             .groups = "drop") %>%
                   left_join(scenarios, by = "scenario") %>%
                   select(scenario = .data$scenario_name, 
                          lake = .data$lake, 
                          metric = .data$metric, 
                          variable = .data$variable, 
                          series = .data$series, 
                          base = .data$base, 
                          sd = .data$sd)


MODFLOW_metrics$scenario <- factor(MODFLOW_metrics$scenario, 
                                   levels = scenarios$scenario_name)

exceeds <- MODFLOW_metrics %>%
           filter(.data$metric == "exceedance_level",
                  .data$scenario == "No Irrigated Agriculture",
                  .data$series == "month") %>%
           mutate(base = NISTmeterTOft(.data$base))
```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Unit Conversions
m_metrics    <- c("exceedance_level", "mean_depth", "max_depth", "median_level",
                  "median_rise_rate", "median_fall_rate")
m2_metrics   <- c("area", "centrarchid_substrate")
m3_metrics   <- c("volume")

# Rounding
d1_metrics   <- c("exceedance_level", "mean_depth", "max_depth", "area",
                  "centrarchid_substrate", "upland_pcnt", "inland_beach_pcnt", 
                  "emergent_pcnt", "floating_pcnt", "submergent_pcnt", 
                  "submergent_algae_pcnt", "submergent_weed_pcnt",
                  "median_level", "cv_level", "median_rise_rate",
                  "median_fall_rate", "season_higher", 
                  "season_lower", "good_spawning")
d0_metrics   <- c("volume", "num_dur", "num_2yr", "median_dur", "cv_dur",
                  "rise_3ft", "rise_1_5ft", "fall_3ft", "fall_1_5ft", 
                  "cv_rise_rate", "cv_fall_rate")

# Friendly names
med_metrics <- data.frame(metric = c("median_level", "cv_level"),
                             name = c("Median Lake Level (ft)",
                                      "CV of Max Depth (%)"))
med_var    <- data.frame(variable = c("0", "1", "4", "7", "10"),
                         var_name = c("Overall", "Winter", "Spring", 
                                      "Summer", "Fall"))

exceed_metrics <- data.frame(metric = c("exceedance_level", 
                                        "area", 
                                        "volume", 
                                        "mean_depth", 
                                        "max_depth"),
                             name = c("Elevation (ft)",
                                      "Area (acres)",
                                      "Volume (acre-ft)",
                                      "Mean Depth (ft)",
                                      "Max Depth (ft)"))
exceed_var <- data.frame(variable = c("10", "25", "50", "75", "90"),
                         var_name = c("Infrequent High", 
                                      "Frequent High",
                                      "Median", 
                                      "Frequent Low",
                                      "Infrequent Low"))

plant_metrics  <- data.frame(metric = c("upland_pcnt", 
                                        "inland_beach_pcnt", 
                                        "emergent_pcnt", 
                                        "floating_pcnt", 
                                        "submergent_pcnt", 
                                        "submergent_algae_pcnt", 
                                        "submergent_weed_pcnt"),
                             name = c("Upland (% of lake)",
                                      "Inland Beach (% of lake)",
                                      "Emergents (% of lake)",
                                      "Floating-Leaf (% of lake)",
                                      "Submergents (% of lake)",
                                      "Submergent Macroalgae (% of lake)",
                                      "Submergent Pondweeds (% of lake)"))
plant_var <- data.frame(variable = c("10", "25", "50", "75", "90"),
                         var_name = c("Infrequent High", 
                                      "Frequent High",
                                      "Median", 
                                      "Frequent Low",
                                      "Infrequent Low"))

dur_metrics <- data.frame(metric = c("num_dur", "num_2yr", 
                                     "median_dur", "cv_dur"),
                             name = c("Times Exceeded for 1+ months",
                                      "Times Exceeded for 2+ yrs",
                                      "Median Duration (months)",
                                      "CV of Duration (%)"))
dur_var    <- data.frame(variable = c("10", "25", "a50", "b50", "75", "90"),
                         var_name = c("Infrequent High", 
                                      "Frequent High",
                                      "Above Median",
                                      "Below Median", 
                                      "Frequent Low",
                                      "Infrequent Low"))

rate_metrics <- data.frame(metric = c("median_rise_rate", 
                                      "cv_rise_rate", 
                                      "rise_1_5ft",
                                      "rise_3ft",
                                      "median_fall_rate", 
                                      "cv_fall_rate", 
                                      "fall_1_5ft",
                                      "fall_3ft"),
                           name = c("Median Rise Rate (ft/time)",
                                    "CV of Rise Rate (%)",
                                    "Times Rise Rate > 1.5 ft/time",
                                    "Times Rise Rate > 3 ft/time",
                                    "Median Fall Rate (ft/time)",
                                    "CV of Fall Rate (%)",
                                    "Times Fall Rate > 1.5 ft/time",
                                    "Times Fall Rate > 3 ft/time"))
rate_var    <- data.frame(variable = c("1", "3", "12"),
                          var_name = c("1 Month", "3 Month", "12 Month"))

time_metrics <- data.frame(metric = c("season_higher", "good_spawning"),
                           name = c("% Years > Prior Season",
                                    "% Years > Prior Growing Season"))
time_var    <- data.frame(variable = c("1", "4", "high_spring", "7", "10"),
                          var_name = c("Winter", "Spring", "Spring", 
                                       "Summer", "Fall"))
```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_levels <- function(df, exceeds, scenarios, lake, 
                        shaded = FALSE, median = FALSE,
                        dur_high = FALSE, dur_low = FALSE,
                        scenario = "No Irrigated Agriculture", text_size = 14) {
  df <- df %>% 
        filter(.data$lake == !!lake,
               .data$scenario == !!scenario,
               .data$sim == 1)
  
  median_line  <- exceeds %>% 
                  filter(.data$lake == !!lake, 
                         .data$variable == "50")
  rectangle   <- exceeds %>% 
                 filter(.data$lake == !!lake,
                        .data$variable %in% c("10", "90")) %>% 
                 select(.data$lake, .data$variable, .data$base) %>%
                 dcast(lake~variable, value.var = "base")
  ribbon_df   <- df %>%
                 mutate(x = .data$time,
                        median = median_line$base,
                        scenario = scenario) %>%
                     select(.data$scenario, .data$x, .data$level, .data$median)
  dur_high_ribbon <- ribbon_df %>%
                     mutate(ymin = .data$median,
                            ymax = ifelse(.data$median > .data$level,
                                          .data$median, .data$level))
  dur_low_ribbon <- ribbon_df %>%
                     mutate(ymax = .data$median,
                            ymin = ifelse(.data$median < .data$level,
                                          .data$median, .data$level))
  
  # Build plot
  plot_obj <- ggplot(df)
  
  # Infrequent High to Low
  if (shaded) {
    plot_obj <- plot_obj +
                geom_rect(data = rectangle,
                          mapping = aes(xmin = -Inf,
                                        xmax = Inf,
                                        ymin = .data$`90`,
                                        ymax = .data$`10`),
                          fill = "steelblue4",
                          color = NA,
                          alpha = 0.1) 
  }
  
  if (median) {
    plot_obj <- plot_obj +
                geom_hline(data = median_line,
                           aes(yintercept = .data$base),
                           linetype = "solid")
  }
  
  if (dur_high) {
    plot_obj <- plot_obj +
                geom_ribbon(data = dur_high_ribbon,
                           aes(x = .data$x,
                               ymin = .data$ymin,
                               ymax = .data$ymax,
                               color = NA,
                               fill = .data$scenario),
                          alpha = 0.1)
  }
  
  if (dur_low) {
    plot_obj <- plot_obj +
                geom_ribbon(data = dur_low_ribbon,
                           aes(x = .data$x,
                               ymin = .data$ymin,
                               ymax = .data$ymax,
                               color = NA,
                               fill = .data$scenario),
                          alpha = 0.1)
  }
  
  plot_obj <- plot_obj +
              geom_line(aes(x = .data$time,
                            y = .data$level,
                            color = .data$scenario),
                        size = 1) +
              scale_color_manual(name = "Scenario",
                                 breaks = scenarios$scenario_name,
                                 values = scenarios$colors,
                                 labels = scenarios$scenario_name) +
              scale_fill_manual(name = "Scenario",
                                 breaks = scenarios$scenario_name,
                                 values = scenarios$colors,
                                 labels = scenarios$scenario_name) +
              labs(x = "Time (yrs)", y = "Elevation (ft)") +
              scale_x_continuous(expand = c(0,0),
                                 breaks = seq(0,34,5),
                                 minor_breaks = seq(0, 34, 1)) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size),
                    legend.position = "none",
                    panel.background = element_rect(fill = "transparent", 
                                                    color = NA),  
                    plot.background = element_rect(fill = "transparent", 
                                                   color = NA))
  
  return(plot_obj)
}
```


### Time series only

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=fig_width, fig.height=3*fig_height}
p1 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Pleasant")
p2 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Long")
p3 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Plainfield")

if (save_on) {
  ggsave("psnt_no_irr.png", p1, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("long_no_irr.png", p2, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("pfl_no_irr.png", p3, 
         device = "png", width = fig_width, height = fig_height, units = "in")
}

p1 + p2 + p3 + plot_layout(ncol = 1, guides = "collect")

```

<br>

### Time series with median level

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=fig_width, fig.height=3*fig_height}
p1 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Pleasant", median = TRUE)
p2 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Long", median = TRUE)
p3 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Plainfield", median = TRUE)

if (save_on) {
  ggsave("psnt_no_irr_median.png", p1, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("long_no_irr_median.png", p2, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("pfl_no_irr_median.png", p3, 
         device = "png", width = fig_width, height = fig_height, units = "in")
}

p1 + p2 + p3 + plot_layout(ncol = 1, guides = "collect")

```

<br>

### Time series with median level and 10%-90% exceedance probability levels shaded
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=fig_width, fig.height=3*fig_height}
p1 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Pleasant", median = TRUE, shaded = TRUE)
p2 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Long", median = TRUE, shaded = TRUE)
p3 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Plainfield", median = TRUE, shaded = TRUE)

if (save_on) {
  ggsave("psnt_no_irr_median_range.png", p1, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("long_no_irr_median_range.png", p2, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("pfl_no_irr_median_range.png", p3, 
         device = "png", width = fig_width, height = fig_height, units = "in")
}

p1 + p2 + p3 + plot_layout(ncol = 1, guides = "collect")

```

<br>

### Time series with duration above the median level shaded

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=fig_width, fig.height=3*fig_height}
p1 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Pleasant", median = TRUE, dur_high = TRUE)
p2 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Long", median = TRUE, dur_high = TRUE)
p3 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Plainfield", median = TRUE, dur_high = TRUE)

if (save_on) {
  ggsave("psnt_no_irr_dur_high.png", p1, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("long_no_irr_dur_high.png", p2, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("pfl_no_irr_dur_high.png", p3, 
         device = "png", width = fig_width, height = fig_height, units = "in")
}

p1 + p2 + p3 + plot_layout(ncol = 1, guides = "collect")

```

<br>

### Time series with duration below the median level shaded

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=fig_width, fig.height=3*fig_height}
p1 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Pleasant", median = TRUE, dur_low = TRUE)
p2 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Long", median = TRUE, dur_low = TRUE)
p3 <- plot_levels(MODFLOW, exceeds, scenarios, lake = "Plainfield", median = TRUE, dur_low = TRUE)

if (save_on) {
  ggsave("psnt_no_irr_dur_low.png", p1, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("long_no_irr_dur_low.png", p2, 
         device = "png", width = fig_width, height = fig_height, units = "in")
  ggsave("pfl_no_irr_dur_low.png", p3, 
         device = "png", width = fig_width, height = fig_height, units = "in")
}

p1 + p2 + p3 + plot_layout(ncol = 1, guides = "collect")

```
