---
title: "MODFLOW_outliers"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MODFLOW_outliers}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSscenarios)
library(dplyr)
library(reshape2)
library(ggplot2)
library(extrafont)
library(lubridate)
library(DT)


rules            <- CSLSscenarios::ecological_rules %>%
                    select(.data$hydrology, .data$metric, .data$variable,
                           .data$Pleasant, .data$Long, .data$Plainfield) %>%
                    unique() %>%
                    melt(id.vars = c("hydrology", "metric", "variable")) 
colnames(rules)  <- c("hydrology", "metric", "variable", "lake", "value")
rules            <- rules %>%
                    filter(.data$value == TRUE) %>%
                    select(.data$lake, .data$hydrology, .data$metric, .data$variable)
useful_metrics   <- unique(rules[,c("lake", "hydrology", "metric", "variable")])
MODFLOW_metrics  <- CSLSscenarios::MODFLOW_metrics %>%
                    filter(.data$series == "month",
                           .data$scenario %in% c("no_irr", "cur_irr")) %>%
                    inner_join(useful_metrics,
                               by = c("lake", "metric", "variable"))

summary_metrics  <- MODFLOW_metrics %>%
                    group_by(.data$lake, .data$scenario,
                             .data$metric, .data$variable) %>%
                    summarise(base = .data$value[which(.data$sim == 1)],
                              q25 = quantile(.data$value,
                                             probs = 0.75,
                                             type = 6,
                                             na.rm = TRUE),
                              q75 = quantile(.data$value,
                                             probs = 0.25,
                                             type = 6,
                                             na.rm = TRUE),
                              IQR = IQR(.data$value, 
                                        type = 6,
                                        na.rm = TRUE),
                              upper = .data$q25 + 1.5*.data$IQR,
                              lower = .data$q75 - 1.5*.data$IQR,
                              sd = sd(.data$value, na.rm = TRUE),
                              .groups = "drop") %>%
                    mutate_at(c("base", "q25", "q75", "IQR", "upper", 
                                "lower", "sd"), as.numeric)

outliers        <- MODFLOW_metrics %>%
                   left_join(summary_metrics, 
                             by = c("lake", "scenario", "metric", "variable")) %>%
                   mutate(outlier = ifelse((.data$value > .data$upper | 
                                              .data$value < .data$lower),
                                           TRUE, FALSE))
outliers$sim     <- factor(outliers$sim)
outliers$outlier <- factor(outliers$outlier)

```

## Overview

This is an exploration of the Monte Carlo SWB simualations, focused on the
no-irrigated-agriculture and the difference between no-irrigated-agriculture and
current-irrigated-agriculture. We tried to use this to understand if any
simulations were outliers, and if so, what caused the differences. Ultimately,
we did not "throw out" any simulations, though we focused primarily on the
"base" parameterization of SWB, as well as the "large drawdown", and "small
drawdown" parameterizations.


I limited this evaluation to only those hydrologic metrics which we use for at
least one of our ecological rules. This is <50% of all metrics we calculate, but
are the metrics most meaningful to us.

<br>

### Plots
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
out1 <- c(304, 233, 25)
out2 <- c(116, 224)


MODFLOW <- CSLSdata::MODFLOW %>%
           filter(year(.data$date) >= 1986,
                  year(.data$date) <= 2018) %>%
           select(.data$scenario, .data$sim, .data$lake, 
                  .data$level_m, .data$date)
MOD1    <- MODFLOW %>% filter(.data$scenario == "no_irr")
MOD2    <- MODFLOW %>% filter(.data$scenario == "cur_irr")
DIFF    <- full_join(MOD1, MOD2,
                     by = c("lake", "sim", "date")) %>%
           mutate(diff = .data$level_m.y - .data$level_m.x)


ggplot() +
  geom_line(data = filter(MOD1, .data$sim %in% c(out1, out2)),
            aes(x = .data$date,
                y = .data$level_m,
                group = as.factor(.data$sim),
                color = as.factor(.data$sim),
                linetype = "No Irr")) +
  geom_line(data = filter(MOD2, .data$sim %in% out1),
            aes(x = .data$date,
                y = .data$level_m,
                group = as.factor(.data$sim),
                color = as.factor(.data$sim),
                linetype = "Current Irr")) +
  scale_linetype_manual(breaks = c("No Irr", "Current Irr"),
                        values = c("solid", "dashed")) +
  scale_color_brewer(name = "Simulation",
                     palette = "Dark2") +
  labs(x = "", y = "Lake Elevation (m)", linetype = "") +
  facet_wrap(~lake, ncol = 1, scales = "free_y") +
  theme_bw()


ggplot() +
  geom_line(data = DIFF,
            aes(x = .data$date,
                y = .data$diff,
                group = as.factor(.data$sim)),
            color = "black",
            alpha = 0.05) +
  geom_line(data = filter(DIFF, .data$sim %in% out1),
            aes(x = .data$date,
                y = .data$diff,
                group = as.factor(.data$sim),
                color = "Reg Outliers")) +
  geom_line(data = filter(DIFF, .data$sim %in% out2),
            aes(x = .data$date,
                y = .data$diff,
                group = as.factor(.data$sim),
                color = "Diff Outliers")) +
  scale_color_manual(name = "Outlier Type",
                     breaks = c("Reg Outliers",
                                "Diff Outliers"),
                     values = c("blue", "red")) +
  labs(x = "", y = "Difference (m)") +
  facet_wrap(~lake, ncol = 1, scales = "free_y") +
  theme_bw()



ggplot() +
  geom_line(data = MOD1,
            aes(x = .data$date,
                y = .data$level_m,
                group = as.factor(.data$sim)),
            color = "black",
            alpha = 0.05) +
  geom_line(data = filter(MOD1, .data$sim %in% c(322)),
            aes(x = .data$date,
                y = .data$level_m,
                group = as.factor(.data$sim)),
            color = "red") +
  labs(x = "", y = "Lake Elevation (m)", linetype = "") +
  facet_wrap(~lake, ncol = 1, scales = "free_y") +
  theme_bw()
```

## Evaluate scenario value independently

I'd consider the **top 3 outliers to be #304, #233, and #25**, but other
simulations still have substantial amounts of outliers for the metrics we care
about.

<br>

### All outliers

Evaluate all metrics used for ecological rules, for all 3 lakes, for all 2
scenarios ("cur_irr" and "no irr"). Total metrics evaluated:

  * 66 for each simulation

<br>

**Top 10 outlier simulations**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

outliers_all <- outliers %>%
                group_by(.data$outlier, .data$sim, .drop = FALSE) %>%
                summarise(count = n(),
                          .groups = "drop") %>%
                filter(.data$outlier == "TRUE") %>%
                arrange(desc(.data$count)) %>%
                select(.data$sim, .data$count)
datatable(outliers_all)

```

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

outliers_all$index <- row.names(outliers_all)
ggplot(outliers_all) +
  geom_point(aes(x = as.numeric(.data$index),
                 y = .data$count)) +
  labs(x = "Simulation Rank",
       y = "Number of Outliers in Simulation") +
  theme_bw()
ggplot(outliers_all) +
  geom_histogram(aes(x = .data$count),
                 binwidth = 1) +
  labs(x = "Number of Outliers",
       y = "Number of Simulations") +
  theme_bw()

```

<br>


### Break down by: hydrology

Break outliers down by type of metric, but lump scenarios and lakes. Total
metrics evaluated:

  * 12 duration
  * 12 frequency
  * 36 magnitude
  * 4 rate of change
  * 2 timing

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
check <- outliers %>% 
         count(.data$sim, 
               .data$hydrology,  
               .drop = FALSE)

outliers_sub <- outliers %>%
                group_by(.data$outlier, .data$sim, 
                         .data$hydrology,  
                         .drop = FALSE) %>%
                summarise(count = n(),
                          .groups = "drop") %>%
                filter(.data$outlier == "TRUE") %>%
                filter(!is.na(.data$hydrology)) %>%
                select(.data$hydrology, .data$sim, .data$count) %>%
                dcast(sim~hydrology, value.var = "count",
                      fill = 0) 
datatable(outliers_sub)

```

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6.5}
plot_outliers_sub <- melt(outliers_sub, id.vars = "sim")
ggplot(plot_outliers_sub) +
  geom_histogram(aes(x = .data$value),
                 binwidth = 1) +
  facet_wrap(~variable) +
  labs(x = "Number of Outliers",
       y = "Number of Simulations") +
  theme_bw()

```

<br>

### Break down by: hydrology and scenario

Break outliers down by type of metric and scenario but lump by lake. Total
metrics evaluated for each scenario:

  * 6 duration
  * 6 frequency
  * 18 magnitude
  * 2 rate of change
  * 1 timing

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
outliers_sub <- outliers %>%
                group_by(.data$outlier, .data$sim, 
                         .data$hydrology, .data$scenario, 
                         .drop = FALSE) %>%
                summarise(count = n(),
                          .groups = "drop") %>%
                filter(.data$outlier == "TRUE") %>%
                filter(!is.na(.data$hydrology)) %>%
                select(.data$hydrology, .data$scenario, .data$sim, .data$count) %>%
                dcast(sim+scenario~hydrology, value.var = "count",
                      fill = 0) 
datatable(outliers_sub)

```

<br>

## Evaluate paired differences

I'd focus on the **top 2 outliers as #224 and #116**, but there's an even
fuzzier line for the top outlier simulations here.

### Outliers for differences 

Evaluate outliers in metric differences (no_irr - cur_irr). Total metrics evaluated:

  * 33 for each simulation
 

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
df1 <- MODFLOW_metrics %>% 
       filter(.data$scenario == "no_irr") %>%
       select(.data$hydrology, .data$metric, .data$variable, .data$lake, 
              .data$sim, .data$value)
df2 <- MODFLOW_metrics %>% 
       filter(.data$scenario == "cur_irr")  %>%
       select(.data$hydrology, .data$metric, .data$variable, .data$lake, 
              .data$sim, .data$value)
combined <- full_join(df1, df2, 
                      by = c("lake", "sim", "metric", "variable", "hydrology"))
combined$diff <- combined$value.x - combined$value.y

summary_combined <- combined %>%
                    group_by(.data$lake, .data$hydrology, .data$metric, 
                             .data$variable) %>%
                    summarise(base = .data$diff[which(.data$sim == 1)],
                              q25 = quantile(.data$diff,
                                             probs = 0.75,
                                             type = 6,
                                             na.rm = TRUE),
                              q75 = quantile(.data$diff,
                                             probs = 0.25,
                                             type = 6,
                                             na.rm = TRUE),
                              IQR = IQR(.data$diff, 
                                        type = 6,
                                        na.rm = TRUE),
                              upper = .data$q25 + 1.5*.data$IQR,
                              lower = .data$q75 - 1.5*.data$IQR,
                              sd = sd(.data$diff, na.rm = TRUE),
                              .groups = "drop") %>%
                    mutate_at(c("base", "q25", "q75", "IQR", "upper", 
                                "lower", "sd"), as.numeric)
outliers        <- combined %>%
                   left_join(summary_combined, 
                             by = c("lake", "hydrology", 
                                    "metric", "variable")) %>%
                   mutate(outlier = ifelse((.data$diff > .data$upper | 
                                              .data$diff < .data$lower),
                                           TRUE, FALSE))
outliers$sim     <- factor(outliers$sim)
outliers$outlier <- factor(outliers$outlier)

outliers_all <- outliers %>%
                group_by(.data$outlier, .data$sim, 
                         .drop = FALSE) %>%
                summarise(count = n(),
                          .groups = "drop") %>%
                filter(.data$outlier == TRUE) %>%
                select(.data$sim, .data$count) %>%
                arrange(desc(.data$count))
datatable(outliers_all)

```


<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

outliers_all$index <- row.names(outliers_all)
ggplot(outliers_all) +
  geom_point(aes(x = as.numeric(.data$index),
                 y = .data$count)) +
  labs(x = "Simulation Rank",
       y = "Number of Outliers in Simulation") +
  theme_bw()
ggplot(outliers_all) +
  geom_histogram(aes(x = .data$count),
                 binwidth = 1) +
  labs(x = "Number of Outliers",
       y = "Number of Simulations") +
  theme_bw()

```

<br>


### Break down by: hydrology

Break outliers down by type of metric, but lump scenarios and lakes. Total
metrics evaluated:

  * 6 duration
  * 6 frequency
  * 18 magnitude
  * 2 rate of change (there are no outliers for rate of change)
  * 1 timing

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
outliers_sub <- outliers %>%
                group_by(.data$outlier, .data$sim, 
                         .data$hydrology,  
                         .drop = FALSE) %>%
                summarise(count = n(),
                          .groups = "drop") %>%
                filter(.data$outlier == "TRUE") %>%
                filter(!is.na(.data$hydrology)) %>%
                select(.data$hydrology, .data$sim, .data$count) %>%
                dcast(sim~hydrology, value.var = "count") 

datatable(outliers_sub)
```

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6.5}
plot_outliers_sub <- melt(outliers_sub, id.vars = "sim")
ggplot(plot_outliers_sub) +
  geom_histogram(aes(x = .data$value),
                 binwidth = 1) +
  facet_wrap(~variable) +
  labs(x = "Number of Outliers",
       y = "Number of Simulations") +
  theme_bw()

```
