---
title: "MODFLOW_comparison"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MODFLOW_comparison}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(CSLSscenarios)
library(reshape2)
library(NISTunits)
library(ggplot2)
library(extrafont)
library(patchwork)
library(lubridate)

MODFLOW         <- CSLSdata::MODFLOW %>%
                   filter(.data$scenario %in% c("no_irr", "cur_irr"),
                          year(.data$date) >= 1986) %>%
                   select(.data$lake, .data$scenario, .data$sim,
                          .data$date, .data$level_m) 
MODFLOW_metrics <- CSLSscenarios::MODFLOW_metrics %>%
                   filter(.data$scenario %in% c("no_irr", "cur_irr")) %>%
                   mutate(value = ifelse(.data$metric ==  "area",
                                         NISTsqrMeterTOacre(.data$value),
                                         ifelse(.data$metric == "volume",
                                                NISTcubMeterTOacreFt(.data$value),
                                                .data$value)))
MODFLOW_diff    <- dcast(MODFLOW_metrics,
                         lake+metric+variable+sim+series~scenario,
                         value.var = "value") %>%
                   mutate(value = .data$cur_irr - .data$no_irr) 


compare_width  <- 6.5
compare_height <- 6
diff_width     <- 6.5
diff_height    <- 5

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_comparison <- function(df, 
                            metric = data.frame(metric = "exceedance_level",
                                                ylabel = "Lake Elevation (ft)",
                                                mTOft = TRUE),
                            variables = data.frame(breaks = c("10", "25", "50",
                                                              "75", "90"),
                                                   labels = c("Infrequent\nHigh",
                                                              "Frequent\nHigh",
                                                              "Median",
                                                              "Frequent\nLow",
                                                              "Infrequent\nLow")),
                            series = "month",
                            colors = data.frame(breaks = c("no_irr", "cur_irr"),
                                                labels = c("No Irrigation",
                                                           "Current Irrigation"),
                                                values = c("darkblue",
                                                           "gold")),
                            text_size = 12,
                            legend_position = "top") {
  
  # Manipulate data frame for plotting
  df         <- df %>%
                filter(.data$metric == !!metric$metric,
                       .data$series == !!series,
                       .data$variable %in% variables$breaks) 
  df$scenario <- factor(df$scenario, levels = colors$breaks)
  df$variable <- factor(df$variable, levels = variables$breaks)
  if (metric$mTOft) { 
    df$value  <- NISTmeterTOft(df$value) 
  }
  
  # Plot
  plot_obj <- ggplot(df) +
              geom_boxplot(aes(x = as.factor(.data$variable),
                               y = .data$value,
                               fill = .data$scenario),
                           position = position_dodge()) +
              scale_fill_manual(name = "",
                                breaks = colors$breaks,
                                labels = colors$labels,
                                values = colors$values) +
              scale_x_discrete(breaks = variables$breaks,
                               labels = variables$labels) +
              labs(x = "", y = metric$ylabel) +
              facet_wrap(~lake, scales = "free_y", ncol = 1) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size),
                    legend.position = legend_position)
  
  return(plot_obj)
}

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_diff <- function(df, 
                      metric = data.frame(metric = "exceedance_level",
                                          ylabel = "Lake Elevation (ft)",
                                          mTOft = TRUE),
                      variables = data.frame(breaks = c("10", "25", "50",
                                                        "75", "90"),
                                             labels = c("Infrequent\nHigh",
                                                        "Frequent\nHigh",
                                                        "Median",
                                                        "Frequent\nLow",
                                                        "Infrequent\nLow")),
                      series = "month",
                      colors = "grey40",
                      text_size = 12) {
  
  # Manipulate data frame for plotting
  df         <- df %>%
                filter(.data$metric == !!metric$metric,
                       .data$series == !!series,
                       .data$variable %in% variables$breaks) 
  df$variable <- factor(df$variable, levels = variables$breaks)
  if (metric$mTOft) { 
    df$value  <- NISTmeterTOft(df$value) 
  }
  
  # Plot
  plot_obj <- ggplot(df) +
              geom_hline(yintercept = 0,
                         color = "black",
                         linetype = "dashed") +
              geom_boxplot(aes(x = as.factor(.data$variable),
                               y = .data$value),
                           fill = colors) +
              scale_x_discrete(breaks = variables$breaks,
                               labels = variables$labels) +
              labs(x = "", y = metric$ylabel) +
              facet_wrap(~lake, ncol = 1) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size))
  
  return(plot_obj)
}

```


## Overview

This document compares the full suite of no-irrigated-agriculture simulations to
the current-irrigated-agriculture simulations. These Monte Carlo simulations
were run to capture the uncertainty in the Soil Water Balance parameterization,
and thus uncertainty in our estimation of recharge. Here, pairwise differences
indicate the difference between a no-irrigated-agriculture simulation and the
corresponding current-irrigated-agriculture simulation with the same SWB
parameterization (i.e., an apples to apples comparison).

## Magnitude

### Exceedance Probabilities

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=compare_width, fig.height=compare_height}
plot_comparison(MODFLOW_metrics)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=diff_width, fig.height=diff_height}
plot_diff(MODFLOW_diff)
```

<br>

### Exceedance Ranges

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=6}
metric <- data.frame(metric = "exceedance_range",
                     ylabel = "Range (ft)",
                     mTOft = TRUE)
variables<- data.frame(breaks = c("range_10_90",
                                  "range_25_75"),
                       labels = c("Infrequent High -\nInfrequent Low",
                                  "Frequent High -\nFrequent Low"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=5}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

### Max Lake Depth

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=compare_width, fig.height=compare_height}
metric <- data.frame(metric = "max_depth",
                     ylabel = "Maximum Depth (ft)",
                     mTOft = TRUE)
variables<- data.frame(breaks = c("10", "25", "50",
                                  "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=diff_width, fig.height=diff_height}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

### Mean Lake Depth

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=compare_width, fig.height=compare_height}
metric <- data.frame(metric = "mean_depth",
                     ylabel = "Mean Depth (ft)",
                     mTOft = TRUE)
variables<- data.frame(breaks = c("10", "25", "50",
                                  "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=diff_width, fig.height=diff_height}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

### Lake Area

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=compare_width, fig.height=compare_height}
metric <- data.frame(metric = "area",
                     ylabel = "Lake Area (ac)",
                     mTOft = FALSE)
variables<- data.frame(breaks = c("10", "25", "50",
                                  "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=diff_width, fig.height=diff_height}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

### Lake Volume

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=compare_width, fig.height=compare_height}
metric <- data.frame(metric = "volume",
                     ylabel = "Lake Volume (ac-ft)",
                     mTOft = FALSE)
variables<- data.frame(breaks = c("10", "25", "50",
                                  "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=diff_width, fig.height=diff_height}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>


### Seasonal Medians

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=6.5}
metric    <- data.frame(metric = "median_level",
                        ylabel = "Median Lake Elevation (ft)",
                        mTOft = TRUE)
variables <- data.frame(breaks = c("1", "4", "7", "10"),
                        labels = c("Winter", "Spring", "Summer", "Fall"))
plot_comparison(MODFLOW_metrics, metric = metric, variables = variables, 
                series = "season")
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
plot_diff(MODFLOW_diff, metric = metric, variables = variables, 
          series = "season")
```

<br>

## Frequency

### Long Lake - Is Lake & Paddle Sports

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=4}
metric    <- data.frame(metric = "is_lake",
                        ylabel = "Percent of Time",
                        mTOft = FALSE)
variables <- data.frame(breaks = c("percent_epa_lake", "percent_lake_warm", 
                                   "percent_open_lake"),
                        labels = c("Is Lake", "Paddle Sport\nPossible", "Is Open\nLake"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3, fig.height=2.5}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

### Move Docks

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=4}
metric    <- data.frame(metric = "dock",
                        ylabel = "Percent of Years",
                        mTOft = FALSE)
variables <- data.frame(breaks = c("percent_install",
                                   "percent_no_move",
                                   "percent_good_year"),
                        labels = c("Install Dock",
                                   "No Move Dock",
                                   "Good Dock"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```

<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=4}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

## Duration

### Number of Times > 2 yrs
Both scenarios are evaluated relative to the "No Irrigation" median levels.

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=6}
metric    <- data.frame(metric = "num_2yr",
                        ylabel = "Number of Times in 33 years",
                        mTOft = FALSE)
variables <- data.frame(breaks = c("a50",
                                   "b50"),
                        labels = c("Above Median",
                                   "Below Median"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```
<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=4}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>


### Median Duration
Both scenarios are evaluated relative to the "No Irrigation" exceedance
probability levels.

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
metric    <- data.frame(metric = "median_dur",
                        ylabel = "Median Duration (months)",
                        mTOft = FALSE)
variables <- data.frame(breaks = c("10", "25", "a50", "b50", "75", "90"),
                        labels = c("Infrequent\nHigh",
                                   "Frequent\nHigh",
                                   "Above\nMedian",
                                   "Below\nMedian",
                                   "Frequent\nLow",
                                   "Infrequent\nLow"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```
<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=5}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

## Rate of Change

### Median Rise Rate

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=6}
metric    <- data.frame(metric = "median_rise_rate",
                        ylabel = "Median Rise Rate (ft/time period)",
                        mTOft = TRUE)
variables <- data.frame(breaks = c("1", "3", "12"),
                        labels = c("1 Month",
                                   "3 Months",
                                   "12 Months"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```
<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=5}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

### Median Fall Rate

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=6}
metric    <- data.frame(metric = "median_fall_rate",
                        ylabel = "Median Fall Rate (ft/time period)",
                        mTOft = TRUE)
variables <- data.frame(breaks = c("1", "3", "12"),
                        labels = c("1 Month",
                                   "3 Months",
                                   "12 Months"))

plot_comparison(MODFLOW_metrics, metric = metric, variables = variables)
```
<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4.5, fig.height=5}
plot_diff(MODFLOW_diff, metric = metric, variables = variables)
```

<br>

## Timing

### Good Spawning Years

**Comparison**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
metric    <- data.frame(metric = "good_spawning",
                        ylabel = "Good Spawning Years (%)",
                        mTOft = FALSE)
variables <- data.frame(breaks = c("high_spring", 
                                   "steady_summer", 
                                   "good_spawning"),
                        labels = c("High Spring",
                                   "Steady Summer",
                                   "Good Spawning"))

plot_comparison(filter(MODFLOW_metrics, .data$lake == "Pleasant"), 
                       metric = metric, variables = variables)
```
<br>

**Pairwise Differences**

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=2.5}
plot_diff(filter(MODFLOW_diff, .data$lake == "Pleasant"),
          metric = metric, variables = variables)
```

<br>
