---
title: "MODFLOW_impact"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MODFLOW_impact}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSscenarios)
library(dplyr)
library(reshape2)
library(ggplot2)
library(extrafont)
library(knitr) # for kable

MODFLOW_comparison <- CSLSscenarios::MODFLOW_comparison

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Most limiting by ecological category -----------------------------------------
limiting      <- MODFLOW_comparison %>%
                 filter(.data$scenario == "cur_irr") %>%
                 group_by(.data$sim_type,
                          .data$lake,
                          .data$hydrology,
                          .data$metric,
                          .data$variable,
                          .data$value1,
                          .data$value2,
                          .data$diff,
                          .data$category,
                          .data$significant_if) %>%
                 mutate(limit_threshold = ifelse(.data$significant_if == "lower",
                                                 max(.data$threshold, na.rm = TRUE),
                                                 min(.data$threshold, na.rm = TRUE))) %>%
                 filter(.data$limit_threshold == .data$threshold) %>%
                 ungroup() %>%
                 select(.data$sim_type,
                        .data$lake,
                        .data$hydrology,
                        .data$metric,
                        .data$variable,
                        .data$category,
                        .data$indicator,
                        .data$impacted,
                        .data$significant_if,
                        .data$value1,
                        .data$threshold,
                        .data$value2,
                        .data$threshold_diff,
                        .data$diff,
                        .data$bathy_significant_if,
                        .data$bathy1,
                        .data$bathy_threshold,
                        .data$bathy2,
                        .data$bathy_threshold_diff,
                        .data$bathy_diff)

# Most limiting by hydrologic metric -------------------------------------------
most_limiting <- MODFLOW_comparison %>%
                 filter(.data$scenario == "cur_irr") %>%
                 group_by(.data$sim_type,
                          .data$lake,
                          .data$hydrology,
                          .data$metric,
                          .data$variable,
                          .data$value1,
                          .data$value2,
                          .data$diff,
                          .data$significant_if) %>%
                 mutate(limit_threshold = ifelse(.data$significant_if == "lower",
                                                 max(.data$threshold, na.rm = TRUE),
                                                 min(.data$threshold, na.rm = TRUE))) %>%
                 filter(.data$limit_threshold == .data$threshold) %>%
                 ungroup() %>%
                 select(.data$sim_type,
                        .data$lake,
                        .data$hydrology,
                        .data$metric,
                        .data$variable,
                        .data$category,
                        .data$indicator,
                        .data$impacted,
                        .data$significant_if,
                        .data$value1,
                        .data$threshold,
                        .data$value2,
                        .data$threshold_diff,
                        .data$diff,
                        .data$bathy_significant_if,
                        .data$bathy1,
                        .data$bathy_threshold,
                        .data$bathy2,
                        .data$bathy_threshold_diff,
                        .data$bathy_diff)

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_diff <- function(df,
                      metric = data.frame(metric = "exceedance_level",
                                          ylabel = "Change in Lake Elevation (ft)"),
                      variables = data.frame(breaks = c("10", "25", "50",
                                                        "75", "90"),
                                             labels = c("Infrequent\nHigh",
                                                        "Frequent\nHigh",
                                                        "Median",
                                                        "Frequent\nLow",
                                                        "Infrequent\nLow")),
                      plot_diff = TRUE,
                      text_size = 10) {
  if (plot_diff) {
    plot_df <- df %>% 
               filter(.data$metric %in% !!metric$metric,
                      .data$variable %in% variables$breaks) %>%
               select(sim_type = .data$sim_type, 
                      lake = .data$lake, 
                      metric_variable = .data$variable, 
                      impacted = .data$impacted,
                      Threshold = .data$threshold_diff,
                      Comparison = .data$diff) %>%
               melt(id.vars = c("sim_type", "lake", "metric_variable", "impacted"))
    scenario_breaks <- c("Threshold", "Comparison")
    scenario_values <- c("grey80", "grey20")
    plot_obj <- ggplot(plot_df) +
                geom_hline(yintercept = 0,
                           color = "black",
                           linetype = "dashed",
                           size = 1)
  } else {
    plot_df <- df %>% 
               filter(.data$metric %in% !!metric$metric,
                      .data$variable %in% variables$breaks) %>%
               select(sim_type = .data$sim_type, 
                      lake = .data$lake, 
                      metric_variable = .data$variable, 
                      impacted = .data$impacted,
                      Reference = .data$value1,
                      Threshold = .data$threshold,
                      Comparison = .data$value2) %>%
               melt(id.vars = c("sim_type", "lake", "metric_variable", "impacted"))
    scenario_breaks <- c("Reference", "Threshold", "Comparison")
    scenario_values <- c("grey90", "grey50", "grey10")
    plot_obj <- ggplot(plot_df)
  }
  
  plot_obj <- plot_obj +
              geom_point(aes(x = .data$metric_variable,
                             y = .data$value,
                             group = .data$variable,
                             shape = .data$sim_type,
                             fill = .data$variable,
                             color = .data$impacted),
                         position = position_dodge(width = 0.5),
                         size = 2.5) +
              scale_shape_manual(name = "Simulation",
                                 breaks = c("permissive", "base", "conservative"),
                                 labels = c("Permissive", "Base", "Conservative"),
                                 values = c(24, 21, 25))  +
              scale_color_manual(name = "Impacted",
                                 breaks = c(TRUE, FALSE),
                                 values = c("red", "darkgreen")) +
              scale_fill_manual(name = "Scenario",
                                 breaks = scenario_breaks,
                                 values = scenario_values) +
              scale_x_discrete(breaks = variables$breaks,
                               labels = variables$labels) +
              guides(color = guide_legend(override.aes = list(shape = 21)),
                     fill = guide_legend(override.aes = list(shape = 21))) +
              labs(x = "", y = metric$ylabel) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size))
  return(plot_obj)
}

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# FUNCTIONS TO CONVERT DATA FRAME TO TABLE
# Function 1/3: round values to desired precision and save as string -----------
round_to_string <- function(val, metric) {
  # Reference
  digits2 <- c("exceedance_level", "exceedance_range", "solute_budget")
  digits1 <- c("stratification", "turtle_bay", "is_lake", "motorboat", 
               "dock", "good_spawning", "paddlesports")
  digits0 <- c("num_2yr")
  
  val <- as.numeric(val)
  
  # Evaluate
  if (metric %in% digits2) {
    string <- sprintf("%.2f", val)
  } else if (metric %in% digits1) {
    string <- sprintf("%.1f", val)
  } else if (metric %in% digits0) {
    string <- sprintf("%.0f", val)
  } else {
    warning("metric not known, add to list")
  }
  
  return(string)
}

# Function 2/3: aggregate when there is a higher and lower threshold -----------
aggregate_threshold <- function(x) {
  if (length(x) == 2) {
    xmin   <- x[which.min(as.numeric(x))]
    xmax   <- x[which.max(as.numeric(x))]
    string <- sprintf("%s-%s",xmin, xmax)
  } else {
    string <- as.character(x)
  }
  return(string)
}

# Function 3/3: format table ---------------------------------------------------
format_table <- function(df,
                         value_var = "threshold",
                         hydrology_order = c("magnitude", "frequency", 
                                             "duration", "rate_of_change", 
                                             "timing", "fluxes"),
                         metric_order = c("exceedance_level", 
                                          "exceedance_range", "stratification", 
                                          "turtle_bay", "motorboat", "paddlesports",
                                          "num_2yr", 
                                          "dock", "is_lake", "good_spawning", 
                                          "solute_budget")) {
  
  # Retain only "base" values + count of impacts
  vals_df   <- df %>% 
               group_by(.data$hydrology, .data$metric, .data$variable, 
                        .data$category, .data$significant_if) %>%
               mutate(impacted = sum(.data$impacted, na.rm = TRUE)) %>%
               ungroup() %>%
               filter(.data$sim_type == "base",
                      !is.na(.data$impacted)) %>%
               select(.data$hydrology, .data$metric, .data$variable, 
                      .data$category, .data$significant_if, .data$impacted, 
                      .data$value1, .data$threshold, .data$value2, 
                      .data$threshold_diff, .data$diff)
  
  # Round values and convert to strings
  string_df <- vals_df %>%
               mutate_at(c("value1", "value2", "threshold", 
                           "threshold_diff", "diff"), 
                         as.character)
  for (i in 1:nrow(vals_df)) {
    for (col_name in c("value1", "value2", "threshold", "threshold_diff", 
                      "diff")) {
      string_df[i,col_name] <- round_to_string(vals_df[i,col_name],
                                               vals_df$metric[i])
    }
  }
  
  # Reshape table
  if (value_var == "threshold") {
    final_cols <- c("hydrology", "metric", "variable", "value1",
                    unique(string_df$category), "value2")
    table_df <- string_df %>%
                select(.data$hydrology, .data$metric, .data$variable,
                       .data$value1, .data$value2, .data$category,
                       .data$threshold) %>%
                dcast(hydrology+metric+variable+value1+value2~category,
                      value.var = value_var, 
                      fun.aggregate = aggregate_threshold,
                      fill = "") %>%
                select(final_cols) %>%
                mutate(hydrology = factor(.data$hydrology, 
                                          levels = hydrology_order),
                       metric = factor(.data$metric, 
                                          levels = metric_order)) %>%
                arrange(.data$hydrology, .data$metric)
  } else if (value_var == "impacted") {
    final_cols <- c("hydrology", "metric", "variable", "value1",
                    unique(string_df$category), "value2")
    table_df <- string_df %>%
                select(.data$hydrology, .data$metric, .data$variable,
                       .data$value1, .data$value2, .data$category,
                       .data$impacted) %>%
                dcast(hydrology+metric+variable+value1+value2~category,
                      value.var = value_var, 
                      fun.aggregate = max) %>%
                mutate_if(is.numeric, function(x) ifelse(is.infinite(x), 
                                                         NA, x)) %>%
                select(final_cols) %>%
                mutate(hydrology = factor(.data$hydrology, 
                                          levels = hydrology_order),
                       metric = factor(.data$metric, 
                                          levels = metric_order)) %>%
                arrange(.data$hydrology, .data$metric)
  } else if (value_var == "threshold_diff") {
    final_cols <- c("hydrology", "metric", "variable", "value1",
                    unique(string_df$category), "diff")
    table_df <- string_df %>%
                select(.data$hydrology, .data$metric, .data$variable,
                       .data$value1, .data$diff, .data$category,
                       .data$threshold_diff) %>%
                dcast(hydrology+metric+variable+value1+diff~category,
                      value.var = value_var, 
                      fun.aggregate = aggregate_threshold,
                      fill = "") %>%
                select(final_cols) %>%
                mutate(hydrology = factor(.data$hydrology, 
                                          levels = hydrology_order),
                       metric = factor(.data$metric, 
                                          levels = metric_order)) %>%
                arrange(.data$hydrology, .data$metric)
  }
  
  return(table_df)
}

```


## Pleasant Lake
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
df <- limiting %>% filter(.data$lake == "Pleasant")
kable(format_table(df, value_var = "threshold"), 
      caption = "Base Thresholds")
kable(format_table(df, value_var = "threshold_diff"), 
      caption = "Base Threshold Differences")
kable(format_table(df, value_var = "impacted"), 
      caption = "Impacted Counts (min 0 to max 3)")

df <- most_limiting %>% filter(.data$lake == "Pleasant")
```

### Exceedance Levels

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.75}
metric    <- data.frame(metric = "exceedance_level",
                        ylabel = "Lake Elevation (ft)")
variables <- data.frame(breaks = c("10", "25", "50", "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)

```

<br>

### Exceedance Ranges

Ugh. This one is weird because evaluates both higher and lower.

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3, fig.height=3.75}
metric    <- data.frame(metric = "exceedance_range",
                        ylabel = "Range (ft)")
variables <- data.frame(breaks = c("range_10_90"),
                        labels = c("Infrequent High -\nInfrequent Low"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```


<br>

### Frequency

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.75}
metric   <- data.frame(metric = c("turtle_bay", "stratification", 
                                  "motorboat", "dock"),
                       ylabel = "Frequency (% of Time)")
variables <- data.frame(breaks = c("percent_connect_warm", "percent_strat", 
                                   "percent_ok", "percent_install", 
                                   "percent_good_year"),
                        labels = c("Turtle Bay\nConnected",
                                   "Stratified\nLake",
                                   "Legal Motorboat\nArea",
                                   "Good Dock\nInstall",
                                   "Good Dock\nYear"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```


<br>

### Duration

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=3.75}
metric   <- data.frame(metric = c("num_2yr"),
                       ylabel = "Number of times")
variables <- data.frame(breaks = c("a50", "b50"),
                        labels = c("2+ yrs\nAbove Median",
                                   "2+ yrs\nBelow Median"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```


<br>

### Rate of Change

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=3.75}
metric   <- data.frame(metric = c("dock"),
                       ylabel = "Percent of Years")
variables <- data.frame(breaks = c("percent_no_move"),
                        labels = c("Dock Not Moved"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```


<br>

### Timing

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=3.75}
metric   <- data.frame(metric = c("good_spawning"),
                       ylabel = "Percent of Years")
variables <- data.frame(breaks = c("good_spawning"),
                        labels = c("Good Spawning\nConditions"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```


<br>


## Long Lake
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
df <- limiting %>% filter(.data$lake == "Long")
kable(format_table(df, value_var = "threshold"), 
      caption = "Base Thresholds")
kable(format_table(df, value_var = "threshold_diff"), 
      caption = "Base Threshold Differences")
kable(format_table(df, value_var = "impacted"), 
      caption = "Impacted Counts (min 0 to max 3)")

df <- most_limiting %>% filter(.data$lake == "Long")
```

### Exceedance Levels

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.75}
metric    <- data.frame(metric = "exceedance_level",
                        ylabel = "Lake Elevation (ft)")
variables <- data.frame(breaks = c("10", "25", "50", "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)

```

<br>

### Exceedance Ranges

Ugh. This one is weird because evaluates both higher and lower.

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3, fig.height=3.75}
metric    <- data.frame(metric = "exceedance_range",
                        ylabel = "Range (ft)")
variables <- data.frame(breaks = c("range_10_90"),
                        labels = c("Infrequent High -\nInfrequent Low"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```

<br>

### Frequency

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.75}
metric   <- data.frame(metric = c("dock", "paddlesports", "is_lake"),
                       ylabel = "Frequency (% of Time)")
variables <- data.frame(breaks = c("percent_install",
                                   "percent_good_year",
                                   "percent_paddle",
                                   "percent_lake",
                                   "percent_open_lake"),
                        labels = c("Good Dock\nInstall",
                                   "Good Dock\nYear",
                                   "Paddle Sports\nPossible",
                                   "Is Lake",
                                   "Is Open\nLake"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```

<br>

### Duration

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=3.75}
metric   <- data.frame(metric = c("num_2yr"),
                       ylabel = "Number of times")
variables <- data.frame(breaks = c("a50", "b50"),
                        labels = c("2+ yrs\nAbove Median",
                                   "2+ yrs\nBelow Median"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```

<br>

### Rate of Change

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=3.75}
metric   <- data.frame(metric = c("dock"),
                       ylabel = "Percent of Years")
variables <- data.frame(breaks = c("percent_no_move"),
                        labels = c("Dock Not Moved"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```

<br>


## Plainfield Lake
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
df <- limiting %>% filter(.data$lake == "Plainfield")
kable(format_table(df, value_var = "threshold"), 
      caption = "Base Thresholds")
kable(format_table(df, value_var = "threshold_diff"), 
      caption = "Base Threshold Differences")
kable(format_table(df, value_var = "impacted"), 
      caption = "Impacted Counts (min 0 to max 3)")

df <- most_limiting %>% filter(.data$lake == "Plainfield")
```

### Exceedance Levels

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.75}
metric    <- data.frame(metric = "exceedance_level",
                        ylabel = "Lake Elevation (ft)")
variables <- data.frame(breaks = c("10", "25", "50", "75", "90"),
                       labels = c("Infrequent\nHigh",
                                  "Frequent\nHigh",
                                  "Median",
                                  "Frequent\nLow",
                                  "Infrequent\nLow"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)

```

<br>

### Exceedance Ranges

Ugh. This one is weird because evaluates both higher and lower.

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3, fig.height=3.75}
metric    <- data.frame(metric = "exceedance_range",
                        ylabel = "Range (ft)")
variables <- data.frame(breaks = c("range_10_90"),
                        labels = c("Infrequent High -\nInfrequent Low"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```

<br>

### Duration

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=3.75}
metric   <- data.frame(metric = c("num_2yr"),
                       ylabel = "Number of times")
variables <- data.frame(breaks = c("a50", "b50"),
                        labels = c("2+ yrs\nAbove Median",
                                   "2+ yrs\nBelow Median"))

plot_diff(df, metric, variables, plot_diff = FALSE)
plot_diff(df, metric, variables, plot_diff = TRUE)
```

<br>
