---
title: "Area Contributing to Impact"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MODFLOW_contributing_area}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSscenarios)
library(NISTunits)
library(dplyr)
library(reshape2)
library(ggplot2)
library(extrafont)
library(knitr) # for kable
library(patchwork)

MODFLOW_comparison <- CSLSscenarios::MODFLOW_comparison  %>%
                      filter(.data$sim_type == "base",
                             !.data$metric %in% c("turtle_bay", "num_2yr"),
                             (.data$hydrology != "frequency" | 
                                .data$metric != "stratification"),
                             .data$sim != 0) %>%
                      mutate(sim = ifelse(.data$scenario == "cur_irr",
                                          0, .data$sim),
                             scenario = ifelse(.data$scenario == "cur_irr",
                                               "wells_off", .data$scenario),
                             threshold_diff = ifelse(.data$variable == "median",
                                                     .data$threshold - .data$value1,
                                                     .data$threshold_diff),
                             diff = ifelse(.data$variable == "median",
                                                     .data$value2 - .data$value1,
                                                     .data$diff)) %>%
                      filter(.data$scenario == "wells_off")

floating_Long_low <- MODFLOW_comparison %>%
                     filter(.data$lake == "Long",
                            .data$variable == "90") %>%
                     select(.data$lake, .data$hydrology, .data$metric, 
                            .data$variable, .data$sim, .data$scenario,
                            .data$sim_type) %>%
                     unique() %>%
                     mutate(category = "plants",
                            indicator = "floating",
                            significant_if = "lower",
                            value1 = 0,
                            value2 = 0,
                            threshold = 0,
                            threshold_diff = 0,
                            diff = 0,
                            bathy1 = 0,
                            bathy_threshold = 0,
                            bathy2 = 0,
                            bathy_threshold_diff = 0,
                            bathy_diff = 0, 
                            bathy_significant_if = "lower")

MODFLOW_comparison <- bind_rows(MODFLOW_comparison, floating_Long_low)


well_rank_dist   <- CSLSscenarios::well_rank_dist
ecological_rules <- CSLSscenarios::ecological_rules %>%
                    select(.data$metric, .data$indicator, .data$bathy_metric)

MODFLOW_comparison <- left_join(MODFLOW_comparison, well_rank_dist,
                                by = c("lake", "sim")) %>%
                      mutate(dist_m = ifelse(is.na(.data$dist_m),
                                             0, .data$dist_m))

variables <- data.frame(variable = c("90", "75", "50", "25", "10", "range_10_90",
                                     "percent_install", "percent_no_move",
                                     "percent_good_year", "percent_ok", 
                                     "percent_paddle", "percent_epa_lake", 
                                     "percent_open_lake", "good_spawning",
                                     "max", "median"),
                        name = c("Infrequent Low (ft)", "Frequent Low (ft)", "Median (ft)", 
                                 "Frequent High (ft)", "Infrequent High (ft)", 
                                 "Range of Levels (ft)", "Can Install Dock (%)", 
                                 "Don't Move Dock (%)", "Good Dock Years (%)", 
                                 "Good Motorboating (%)", "Good Paddlesports (%)", "Is Lake (%)", 
                                 "Is Open Lake (%)", "Good Pike Spawning (%)",
                                 "Max Mg (mg/L)", "Median Mg (mg/L)"))
MODFLOW_comparison <- left_join(MODFLOW_comparison, variables, by = "variable")
MODFLOW_comparison$name <- factor(MODFLOW_comparison$name,
                                  levels = variables$name)

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Most limiting by ecological category -----------------------------------------
summary      <- MODFLOW_comparison %>%
                 group_by(.data$scenario,
                          .data$sim,
                          .data$lake,
                          .data$hydrology,
                          .data$metric,
                          .data$variable,
                          .data$name,
                          .data$value1,
                          .data$value2,
                          .data$diff,
                          .data$category,
                          .data$significant_if) %>%
                 mutate(limit_threshold = ifelse(.data$significant_if == "lower",
                                                 max(.data$threshold, na.rm = TRUE),
                                                 min(.data$threshold, na.rm = TRUE))) %>%
                 filter(.data$limit_threshold == .data$threshold) %>%
                 ungroup() %>%
                 group_by(.data$lake, .data$sim, .data$dist_m, .data$scenario, 
                          .data$metric, .data$variable, .data$name, .data$category) %>%
                 summarise(indicator = ifelse(length(.data$indicator) > 1,
                                              ifelse(max(.data$value1) > max(.data$value2),
                                                     .data$indicator[which(.data$significant_if == "lower")],
                                                     .data$indicator[which(.data$significant_if == "higher")]),
                                              .data$indicator),
                           impacted = max(.data$impacted),
                           value1 = max(.data$value1),
                           value2 = ifelse(length(.data$value2) > 1,
                                           ifelse(max(.data$value1) > max(.data$value2),
                                                  .data$value2[which(.data$significant_if == "lower")],
                                                  .data$value2[which(.data$significant_if == "higher")]),
                                           .data$value2),
                           threshold = ifelse(length(.data$threshold) > 1,
                                              ifelse(max(.data$value1) > max(.data$value2),
                                                     .data$threshold[which(.data$significant_if == "lower")],
                                                     .data$threshold[which(.data$significant_if == "higher")]),
                                              .data$threshold),
                           diff = ifelse(length(.data$diff) > 1,
                                         ifelse(max(.data$value1) > max(.data$value2),
                                                .data$diff[which(.data$significant_if == "lower")],
                                                .data$diff[which(.data$significant_if == "higher")]),
                                         .data$diff),
                           threshold_diff = ifelse(length(.data$threshold_diff) > 1,
                                                   ifelse(max(.data$value1) > max(.data$value2),
                                                          .data$threshold_diff[which(.data$significant_if == "lower")],
                                                          .data$threshold_diff[which(.data$significant_if == "higher")]),
                                                   .data$threshold_diff),
                           bathy1 = max(.data$value1),
                           bathy2 = ifelse(length(.data$bathy2) > 1,
                                           ifelse(max(.data$bathy1) > max(.data$bathy2),
                                                  .data$bathy2[which(.data$bathy_significant_if == "lower")],
                                                  .data$bathy2[which(.data$bathy_significant_if == "higher")]),
                                           .data$bathy2),
                           bathy_threshold = ifelse(length(.data$bathy_threshold) > 1,
                                                    ifelse(max(.data$bathy1) > max(.data$bathy2),
                                                           .data$bathy_threshold[which(.data$bathy_significant_if == "lower")],
                                                           .data$bathy_threshold[which(.data$bathy_significant_if == "higher")]),
                                                    .data$bathy_threshold),
                           bathy_diff = ifelse(length(.data$bathy_diff) > 1,
                                               ifelse(max(.data$bathy1) > max(.data$bathy2),
                                                      .data$bathy_diff[which(.data$bathy_significant_if == "lower")],
                                                      .data$bathy_diff[which(.data$bathy_significant_if == "higher")]),
                                               .data$bathy_diff),
                           bathy_threshold_diff = ifelse(length(.data$bathy_threshold_diff) > 1,
                                                         ifelse(max(.data$bathy1) > max(.data$bathy2),
                                                                .data$bathy_threshold_diff[which(.data$bathy_significant_if == "lower")],
                                                                .data$bathy_threshold_diff[which(.data$bathy_significant_if == "higher")]),
                                                         .data$bathy_threshold_diff),
                           significant_if = ifelse(length(.data$significant_if) > 1,
                                                         ifelse(max(.data$value1) > max(.data$value2),
                                                                .data$significant_if[which(.data$bathy_significant_if == "lower")],
                                                                .data$significant_if[which(.data$bathy_significant_if == "higher")]),
                                                         .data$significant_if),
                           .groups = "drop") %>%
                 mutate(impacted = ifelse(.data$impacted == 1,
                                          TRUE, 
                                          ifelse(.data$impacted == 0, 
                                                 FALSE, NA)))

```

## Background

In order to understand what area contributes to the impact at Long and
Plainfield lakes, we ran a scenario in which we turned wells off and switched
land use to the no-irrigated-agriculture value one at a time. We started with
the well closest to the centroid of Long Lake, then progressively increased the
zone of no-irrigated-agriculture until all of the 321 wells which were simulated
in the inset model had been removed with land use switched to the
no-irrigated-agriculture value.

<br>

## Overall Summary

The figure below shows the number of ecosystem indicators that are classified as
unimpacted as the radius of the no-irrigated-agriculture area surrounding Long
and Plainfield lakes increases.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=2.5}
total_impacts <- summary %>%
                 group_by(.data$lake, .data$scenario, .data$sim, .data$dist_m) %>%
                 summarise(n = n(),
                           impacted = sum(.data$impacted, na.rm = TRUE),
                           not_impacted = .data$n - .data$impacted,
                           .groups = "drop")

plot_df <- total_impacts %>% 
           filter(.data$lake != "Pleasant") %>%
           mutate(lake = sprintf("%s Lake", .data$lake))

stars <- plot_df %>%
         group_by(.data$lake) %>%
         mutate(no_wells = .data$sim[min(which(.data$impacted == 0))]) %>%
         filter(.data$no_wells == .data$sim) %>%
         ungroup()

ggplot(plot_df) +
  geom_line(aes(x = NISTmeterTOmile(.data$dist_m),
                y = .data$not_impacted)) +
  geom_hline(data = unique(plot_df[,c("lake", "n")]),
             aes(yintercept = .data$n),
             linetype = "dashed") +
  geom_point(data = stars,
             aes(x = NISTmeterTOmile(.data$dist_m), 
                 y = .data$not_impacted),
             shape = 17,
             size = 3) +
  geom_text(data = stars, 
           aes(x = NISTmeterTOmile(.data$dist_m) + 1.5, 
               y = .data$not_impacted - 3,
               label = sprintf("%s wells\n%0.1f miles", 
                               .data$sim, 
                               NISTmeterTOmile(.data$dist_m))),
           family = "Segoe UI Semilight") +
  labs(x = "Radius around Long Lake with no irrigated agriculture (miles)",
       y = "Number of\nUnimpacted Indicators") +
  scale_x_continuous(breaks = seq(0,10,2),
                     minor_breaks = seq(0,10,1),
                     expand = c(0,0)) +
  facet_grid(~lake) +
  theme_bw() +
  theme(text= element_text(family = "Segoe UI Semilight",
                           size = 12))
```

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_value_indicators <- function(df, lake) {
  
  thresholds <- df %>%
                filter(.data$lake == !!lake) %>%
                mutate(ymin = ifelse(.data$significant_if == "lower",
                                     .data$threshold_diff, -Inf),
                ymax = ifelse(.data$significant_if == "higher",
                              .data$threshold_diff, Inf)) %>%
                select(.data$name, .data$category, .data$threshold_diff, 
                       .data$ymin, .data$ymax) %>%
                unique()
  
  rectangles <- thresholds %>%
                group_by(.data$name, .data$category) %>%
                summarise(ymin = max(.data$ymin),
                          ymax = min(.data$ymax),
                          .groups = "drop")
  
  plot_obj <- ggplot(filter(summary, .data$lake == !!lake)) +
              geom_hline(yintercept = 0,
                         size = 1) +
              geom_line(aes(x = NISTmeterTOmile(.data$dist_m),
                            y = .data$diff)) +
              geom_hline(data = thresholds,
                         aes(yintercept = .data$threshold_diff,
                             color = .data$category),
                         size = 1)  +
              geom_rect(data = rectangles,
                        aes(xmin = -Inf,
                            xmax = Inf,
                            ymin = .data$ymin,
                            ymax = .data$ymax,
                            fill = .data$category),
                        color = NA,
                        alpha = 0.3) +
              facet_wrap(~name, scales = "free_y", ncol = 5) +
              labs(x = "Radius around Long Lake with no irrigated agriculture (miles)",
                   y = "Difference relative to No Irrigated Agriculture") +
              scale_x_continuous(expand = c(0,0),
                                 breaks = c(0,5,10),
                                 minor_breaks = seq(0,10,1)) +
              scale_color_manual(name = "Significant Thresholds",
                       breaks = c("human_use", "fish", "plants", "chemistry"),
                       labels = c("Human Use", "Fish", "Plants", "Chemistry"),
                       values = c("#E78AC3", "#8DA0CB", "#66C2A5", "#FC8D62")) +
              scale_fill_manual(name = "Significant Thresholds",
                                breaks = c("human_use", "fish", "plants", "chemistry"),
                                labels = c("Human Use", "Fish", "Plants", "Chemistry"),
                                values = c("#E78AC3", "#8DA0CB", "#66C2A5", "#FC8D62"))+
              theme_bw() +
              theme(text= element_text(family = "Segoe UI Semilight",
                                       size = 8),
                    legend.position = "top")
  return(plot_obj)
}

```


```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 6}
bathy_metrics <- left_join(MODFLOW_comparison, ecological_rules,
                           by = c("metric", "indicator")) %>%
                 filter(.data$metric == "exceedance_level",
                        .data$lake != "Pleasant") %>%
                 mutate(bathy_metric = ifelse(is.na(.data$bathy_metric),
                                              .data$indicator, 
                                              .data$bathy_metric)) %>% 
                 select(.data$lake, .data$category, .data$indicator, 
                        .data$bathy_metric, .data$variable, 
                        .data$bathy_threshold_diff, .data$bathy_diff, 
                        .data$bathy_significant_if, .data$sim, .data$dist_m) %>%
                 filter(.data$bathy_metric != "")

bathy_variables <- data.frame(bathy_metric = c("vol_m3", "area_m2", "upland", 
                                               "inland_beach", "emergent", 
                                               "floating", "submergent"),
                              bathy_name = c("Lake Volume\n(ac-ft)", 
                                             "Lake Area\n(ac)", 
                                             "Upland\n(% Lake)",
                                             "Inland Beach\n(% Lake)", 
                                             "Emergents\n(% Lake)",
                                             "Floating Leaf\n(% Lake)",
                                             "Submergents\n(% Lake)"))
level_variables <- data.frame(variable = c("90", "75", "50", "25", "10"),
                              var_name = c("Infrequent Low", "Frequent Low", "Median", 
                                           "Frequent High", "Infrequent High"))
bathy_metrics <- left_join(bathy_metrics, bathy_variables,
                           by = "bathy_metric") %>%
                 left_join(level_variables, by = "variable")
bathy_metrics$var_name <- factor(bathy_metrics$var_name, level_variables$var_name)

plot_mag_thresholds <- function(df, lake, bathy_metric) {
  if (bathy_metric == "area_m2") {
    df <- df %>%
          mutate(bathy_diff = NISTsqrMeterTOacre(.data$bathy_diff),
                 bathy_threshold_diff = NISTsqrMeterTOacre(.data$bathy_threshold_diff))
  } else if (bathy_metric == "vol_m3") {
    df <- df %>%
          mutate(bathy_diff = NISTcubMeterTOacreFt(.data$bathy_diff),
                 bathy_threshold_diff = NISTcubMeterTOacreFt(.data$bathy_threshold_diff))
  }
  
  plot_df <- df %>% 
             filter(.data$lake == !!lake, 
                    bathy_metric == !!bathy_metric)
  thresholds <- unique(plot_df[,c("bathy_threshold_diff", "var_name",
                                  "bathy_name", "category", "bathy_significant_if")]) %>%
                mutate(ymin = ifelse(.data$bathy_significant_if == "lower",
                                     .data$bathy_threshold_diff, -Inf),
                       ymax = ifelse(.data$bathy_significant_if == "higher",
                                     .data$bathy_threshold_diff, Inf))
  plot_obj <- ggplot(plot_df) +
              geom_hline(yintercept = 0,
                         size = 1) +
              geom_line(aes(x = NISTmeterTOmile(.data$dist_m),
                            y = .data$bathy_diff)) +
              geom_hline(data = thresholds,
                         aes(yintercept = .data$bathy_threshold_diff,
                             color = .data$category),
                         size = 1)  +
              geom_rect(data = thresholds,
                        aes(xmin = -Inf,
                            xmax = Inf,
                            ymin = .data$ymin,
                            ymax = .data$ymax,
                            fill = .data$category),
                        color = NA,
                        alpha = 0.3) +
              facet_wrap(~var_name, nrow = 1) +
              scale_color_manual(name = "",
                                 breaks = c("human_use", "fish", "plants", "chemistry"),
                                 labels = c("Human Use", "Fish", "Plants", "Chemistry"),
                                 values = c("#E78AC3", "#8DA0CB", "#66C2A5", "#FC8D62")) +
              scale_fill_manual(name = "",
                                 breaks = c("human_use", "fish", "plants", "chemistry"),
                                 labels = c("Human Use", "Fish", "Plants", "Chemistry"),
                                 values = c("#E78AC3", "#8DA0CB", "#66C2A5", "#FC8D62")) +
              scale_x_continuous(breaks = c(0,5,10),
                                 minor_breaks = seq(0,10,1)) +
              labs(y = unique(plot_df$bathy_name),
                   x = "Radius around Long Lake with no irrigated agriculture (miles)") +
              theme_bw() +
              theme(text= element_text(family = "Segoe UI Semilight",
                                       size = 10),
                    legend.position = "top")
  return(plot_obj)
}

```

## Long Lake


### Ecosystem Indicators

Here, we take a closer look at the severity of impact to each ecosystem
indicator. The plot below shows key hydrologic metrics as well as significant
thresholds (colored lines) and zones where the metric is unimpacted (shaded
areas) for different types of ecosystem indicators (human use, fish, plants, and
chemistry).

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 5}
plot_value_indicators(summary, "Long") 
```

<br>

### Detailed Bathymetry-related Metrics

Here, we zoom in on the exceedance level metrics and break down how different
ecosystem indicators related to bathymetry are impacted.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=8}
p1 <- plot_mag_thresholds(bathy_metrics, "Long", "vol_m3") + xlab("")
p2 <- plot_mag_thresholds(bathy_metrics, "Long", "area_m2") + xlab("")
p3 <- plot_mag_thresholds(bathy_metrics, "Long", "upland") + xlab("")
p4 <- plot_mag_thresholds(bathy_metrics, "Long", "inland_beach") + xlab("")
p5 <- plot_mag_thresholds(bathy_metrics, "Long", "emergent") + xlab("")
p6 <- plot_mag_thresholds(bathy_metrics, "Long", "floating") + xlab("")
p7 <- plot_mag_thresholds(bathy_metrics, "Long", "submergent")

p1 + p2 + p3 + p4 + p5 + p6 + p7 + plot_layout(guides = "collect", ncol = 1) &
  theme(legend.position = "top",
        plot.margin = margin(0,1,0,1))
```

<br>

## Plainfield Lake

### Ecosystem Indicators

Here, we take a closer look at the severity of impact to each ecosystem
indicator. The plot below shows key hydrologic metrics as well as significant
thresholds (colored lines) and zones where the metric is unimpacted (shaded
areas) for different types of ecosystem indicators (human use, plants, and
chemistry).

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 3.75}
plot_value_indicators(summary, "Plainfield") 
```

<br>

### Detailed Bathymetry-related Metrics
Here, we zoom in on the exceedance level metrics and break down how different
ecosystem indicators related to bathymetry are impacted.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=5.8}
p1 <- plot_mag_thresholds(bathy_metrics, "Plainfield", "upland") + xlab("")
p2 <- plot_mag_thresholds(bathy_metrics, "Plainfield", "inland_beach") + xlab("")
p3 <- plot_mag_thresholds(bathy_metrics, "Plainfield", "emergent") + xlab("")
p4 <- plot_mag_thresholds(bathy_metrics, "Plainfield", "floating") + xlab("")
p5 <- plot_mag_thresholds(bathy_metrics, "Plainfield", "submergent")

p1 + p2 + p3 + p4 + p5 + plot_layout(guides = "collect", ncol = 1) &
  theme(legend.position = "top",
        plot.margin = margin(0,1,0,1))
```
